Deep Reinforcement Learning for Conservation Decisions

### 2021-05-18 manuscript/manuscript.Rmd
Some revisions to the formal section

Training a deep, model-free agent on an environment involves allowing the agent to interact with the environment for thousands or even millions of time steps.
The training period involves a trial and error process whereby the agent will perform an action and then receive feedback from the environment in the form of a reward. 
This process will repeat until the environment transitions to a terminal state, after which the environment will reset to an initial state.
According to the algorithm that the agent follows, the agent will explore the state-action space and use its experience to improve its policy.
The amount of time needed for an agent to learn high reward yielding behavior cannot be predetermined and depends on a host of factors including the amount of dedicated compute, complexity of the environment, complexity of the agent and more.
Yet, overall, it has been well established that deep, model-free RL agents tend to be very sample inefficient [@gu_q-prop_2017], so it is recommended to provide a generous training budget for these agents.
An advantage of using deep, model-free algorithms is that since they function autonomously and do not require any knowledge about the environment, we can readily test how these algorithms perform without knowing much about the environment's transition dynamics or all the inner workings of the algorithm.

@@

# RL Background

When the environment is fully observable, i.e. when the state that the agent observes fully characterizes the system, the reinforcement learning problem is typically formalized as a Markov Decision Process (MDP).
A MDP consists of the following: a set of states, $\mathcal{S}$, called the state space; a set of actions, $\mathcal{A}$, called the action space; a state transition operator, $T(s_{t+1}|s_t, a_t)$, which describes the dynamics of the system; a reward function, $r(s_t, a_t)$; an initial state distribution, $d_0(s_0)$; and a discount factor, $\gamma \in (0,1]$.
After the agent observes the state at time t, $s_t$, the agent selects an action, $a_t$, according to its policy, $\pi$, which maps states to a probability distribution over actions, $\pi: \mathcal{S} \rightarrow p(\mathcal{A} \mid \mathcal{S})$.
As the agent interacts with the environment, the agent will create a trajectory in state-action space given as $\tau = (s_0, a_0, \dots, s_H, a_H)$ where H denotes the horizon or length of the state-action sequence.
From these definitions, we can provide an agent's trajectory distribution for a given policy as,

$$
  p_\pi(\tau) = d_0(s_0) \prod_{t=0}^H \pi(a_t \mid s_t) \, T(s_{t+1} \mid s_t, a_t)
$$

When the agent cannot fully observe the state of the system, reinforcement learning can then be described by a Partially Observable Markov Decision Process (POMDP), wherein there is an additional distribution, $E(o_{t}| s_{t})$, which accounts for the observation being different from system's actual state.
An important property of POMDPs and MDPs is that they are memoryless; the future, $s_{t+1}$, does not depend on the past but only on the state and action at the present time.
This Markov property, which is often an unrealistic assumption, is held in most of the RL algorithms that we will discuss.

In formal terms, the goal of reinforcement learning is to find an optimal policy distribution, $\pi^*(a_t | s_t)$ or $\pi^*(a_t|o_t)$ for MDP and POMDP cases respectively, that maximizes the expected return, $J(\pi)$:

@@

-->
```{=latex}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Algorithm & Description & Model & Policy & Action Space & State Space \\
\hline
MBPO & Model-based Policy Optimization [@mbpo] & Model-based & On-policy & Continuous & Continuous\\
\hline
MCTS & Monte Carlo Tree Search [@suttonbarto] & Model-based & Either & & \\
\hline
A2C  & Advantage Actor Critic [@a2c] & Model-free & On-policy& &\\
\hline
A3C  & Asynchronous A2C & Model-free [@a3c] & On-policy & &\\
\hline
TRPO  & Trust Region Policy Optimization [@trpo] & Model-free & On-policy & &\\
\hline
PPO  & Proximal Policy Optimization [@ppo] & Model-free & On-policy & &\\
\hline
DQN & Deep Q Networks [@dqn] & Model-free & Off-policy& &\\
\hline
DDPG & Deep Deterministic Policy Gradient [@ddpg] & Model-free & Off-policy & &\\
\hline
TD3 & Twin Delayed DDPG [@td3] & Model-free & Off-policy & &\\
\hline
SAC & Soft Actor Critic [@sac] & Model-free & Off-policy & &\\
\hline
IMPALA & Importance Weighted Actor Learner [@impala] & Model-free & Off-policy & &\\
\hline
APE-X & Distributed Prioritized Experience Replay [@apex] & Model-free & Off-policy & & \\
\hline
\end{tabular}
\end{center}
```
# Examples

>>>>

Training a deep RL agent on an environment involves allowing the agent to interact with the environment for thousands or even millions of time steps.
The training period involves a trial and error process whereby the agent will perform an action and then receive feedback from the environment in the form of a reward. 
This process will repeat until the environment transitions to a terminal state, after which the environment will reset to an initial state.
According to the algorithm that the agent follows, the agent will explore the state-action space and use its experience to improve its policy.
The amount of time needed for an agent to learn high reward yielding behavior cannot be predetermined and depends on a host of factors including the amount of dedicated compute, complexity of the environment, complexity of the agent and more.
Yet, overall, it has been well established that deep, model-free RL agents tend to be very sample inefficient [@gu_q-prop_2017], so it is recommended to provide a generous training budget for these agents.
An advantage of using deep, model-free algorithms is that since they function autonomously and do not require any knowledge about the environment, we can readily test how these algorithms perform without knowing much about the environment's transition dynamics or all the inner workings of the algorithm.

@@

# Formal RL Concepts

The reinforcement learning problem is typically formalized as a partially observable Markov decision process (POMDP).
A POMDP is a 7-tuple that consists of the following: 
- $\mathcal{S}$: a set of states called the state space
- $\mathcal{A}$: a set of actions called the action space
- $\Omega$ : a set of observations called the  observation space
- $E(o_{t}| s_{t})$: an emission distribution, which accounts for an agent's observation being different from environment's state
- $T(s_{t+1}|s_t, a_t)$: a state transition operator which describes the dynamics of the system
- $r(s_t, a_t)$: a reward function
- $d_0(s_0)$: an initial state distribution
- $\gamma \in (0,1]$: a discount factor

After the agent makes an observation at time t, $o_t$, the agent selects an action, $a_t$, according to its policy, $\pi$, which maps observations to a probability distribution over actions, $\pi: \Omega \rightarrow p(\mathcal{A} \mid \Omega)$.
This action will cause the environment to transition to a new state according to its state transition operator. 
The agent will then receive a new observation and the process will repeat until a terminal state is reached.
As the agent interacts with the environment, the agent will create a trajectory in state-action space given as $\tau = (s_0, a_0, \dots, s_H, a_H)$ where H denotes the horizon or length of the state-action sequence.
From these definitions, we can provide an agent's trajectory distribution for a given policy as,

$$
  p_\pi(\tau) = d_0(s_0) \prod_{t=0}^H \pi(a_t | o_t) \, E(o_{t}| s_{t}) \, T(s_{t+1} | s_t, a_t)
$$

An important property of POMDPs and MDPs is that they are memoryless; the future, $s_{t+1}$, does not depend on the past but only on the state and action at the present time.
This Markov property, which is often an unrealistic assumption, is held in most of the RL algorithms that we discuss in the next section.

In formal terms, the goal of reinforcement learning is to find an optimal policy distribution, $\pi^*(a_t|o_t)$, that maximizes the expected return, $J(\pi)$:

@@

-->
```{=latex}
\begin{center}
\begin{tabular}{|c|c|c|c|c|c|}
\hline
Algorithm & Description & Model & Evaluation & Policy & Constraints  \\
\hline
MBPO & Model-based Policy Optimization [@mbpo] & Model-based & On-policy & Deterministic & \\
\hline
MCTS & Monte Carlo Tree Search [@suttonbarto] & Model-based & Either & & \\
\hline
A2C  & Advantage Actor Critic [@a2c] & Model-free & On-policy& &\\
\hline
A3C  & Asynchronous A2C & Model-free [@a3c] & On-policy & &\\
\hline
TRPO  & Trust Region Policy Optimization [@trpo] & Model-free & On-policy & &\\
\hline
PPO  & Proximal Policy Optimization [@ppo] & Model-free & On-policy & &\\
\hline
DQN & Deep Q Networks [@dqn] & Model-free & Off-policy& &\\
\hline
DDPG & Deep Deterministic Policy Gradient [@ddpg] & Model-free & Off-policy & &\\
\hline
TD3 & Twin Delayed DDPG [@td3] & Model-free & Off-policy & &\\
\hline
SAC & Soft Actor Critic [@sac] & Model-free & Off-policy & &\\
\hline
IMPALA & Importance Weighted Actor Learner [@impala] & Model-free & Off-policy & &\\
\hline
APE-X & Distributed Prioritized Experience Replay [@apex] & Model-free & Off-policy & & \\
\hline
\end{tabular}
\end{center}
```
# Examples
